Robinson et al. 2020 Supplemental Information: Comparing the Radiocarbon and Tree-ring Records for the Upland US Southwest

### 2020-01-19 analysis/paper/paper.Rmd
trying precommit

theme_map <- function(base_size = 6.5,
                      base_family = ""){
  theme_bw(base_size = base_size,
           base_family = base_family) %+replace%
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          panel.background = element_blank(),
          panel.border = element_blank(),
          panel.grid = element_blank(),
          panel.grid.major = element_line(colour = 'transparent'),
          panel.grid.minor = element_line(colour = 'transparent'),
          legend.key = element_blank(),
          legend.position = "top",
          plot.background = element_blank(),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

ggplot() +
  geom_sf(data = four_corners_states) +
  geom_raster(data = robinson2020::swus_hillshade_500m,
              mapping = aes(x = x,
                            y = y,
                            alpha = ID),
              na.rm = TRUE) +
  scale_alpha(range = c(0.8, 0),
              na.value = 0,
              limits = c(0,255),
              guide = "none") +
  geom_sf(data = four_corners_states,
          fill = NA) +
  geom_sf(data = uusw_counties %>% 
            sf::st_union() %>%
            rmapshaper::ms_filter_islands(),
          fill = NA,
          color = "black") +
  geom_sf(data = uusw_boundary,
          fill = NA,
          linetype = 2,
          color = "white") +
  geom_sf(data = vep_study_areas,
          fill = NA,
          color = "white",
          size = 1) +
  geom_sf_text(data = vep_study_areas,
               aes(label = `Study Area`),
               size = 3) +
  theme_map(16)

```

@@

```{r prepare-radiocarbon-data}
# Read in the radiocarbon database
if(!file.exists(
  here::here("analysis/data/derived_data/radiocarbon_data.csv")
))
  readr::read_csv(
    here::here("analysis/data/raw_data/0403_sbox_clean_sd300.csv")
  ) %>%
  # Convert to spatial object
  sf::st_as_sf(coords = c("Long","Lat"),
               crs = 4326) %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  # Select salient columns
  dplyr::select(labnumber,
                date,
                sd,
                Site,
                SiteID) %>%
  dplyr::rename(Lab_Number = labnumber,
                Date_BP = date,
                SD = sd,
                Site_Number = Site) %>%
  sf::st_intersection(uusw_counties) %>%
  sf::st_drop_geometry() %>%
  dplyr::mutate(Type = "Radiocarbon",
                SiteID = SiteID %>% stringr::str_c("Radiocarbon - ", .)) %>%
  dplyr::select(Type,
                SiteID,
                Site_Number,
                County,
                Lab_Number,
                Date_BP,
                SD) %>%
  # Write the output to a new file (csv)
  readr::write_csv(
    here::here("analysis/data/derived_data/radiocarbon_data.csv")
  )

radiocarbon_data <-
  readr::read_csv(
    here::here("analysis/data/derived_data/radiocarbon_data.csv")
  )

```

@@

if(!file.exists(
  here::here("analysis/data/derived_data/dendro_data.csv")
))
  readr::read_csv(
    here::here("analysis/data/raw_data/BOCINSKY_ET_AL_2015_TREE_RINGS.csv")
  ) %>%
  # Convert to spatial object
  sf::st_as_sf(coords = c("LON","LAT"),
               crs = 4326) %>%
  sf::st_transform(four_corners_lcc_proj) %>%
  # Convert dates to BP
  dplyr::mutate(Date_BP = 1950 - Outer_Date_AD,
                Type = ifelse(C_Level %in% c(2,3),
                              "Cutting or Near-cutting",
                              "Non-cutting")) %>%
  # drop C_Level
  dplyr::select(Seq_Number,
                Site_Number,
                Lab_Number,
                Type,
                Date_BP) %>%
  dplyr::rename(SiteID = Seq_Number) %>%
  sf::st_intersection(uusw_counties) %>%
  sf::st_drop_geometry() %>%
  dplyr::mutate(Type = Type %>% stringr::str_c("Dendro - ", .),
                SiteID = SiteID %>% stringr::str_c("Dendro - ", .),
                SD = 0) %>%
  dplyr::filter(Type == "Dendro - Cutting or Near-cutting") %>%
  dplyr::mutate(Type = "Tree-ring")  %>%
  dplyr::select(Type,
                SiteID,
                Site_Number,
                County,
                Lab_Number,
                Date_BP,
                SD) %>%
  # Write the output to a new file (csv)
  readr::write_csv(
    here::here("analysis/data/derived_data/dendro_data.csv")
  )

dendro_data <- 
  readr::read_csv(
    here::here("analysis/data/derived_data/dendro_data.csv")
  )

```

@@

```{r}
all_data <- list(Radiocarbon = radiocarbon_data,
                 `Tree-ring` = dendro_data) %>%
  purrr::map(~dplyr::mutate(.x,
                            `Study Area` = ifelse(County == "Montezuma, CO",
                                                  "CMV", 
                                                  NA),
                            `Study Area` = ifelse(County %in% c("Los Alamos, NM",
                                                                "Rio Arriba, NM",
                                                                "Sandoval, NM",
                                                                "Santa Fe, NM",
                                                                "Taos, NM"),
                                                  "NRG", 
                                                  `Study Area`)))

```

@@

```{r data-summary}
all_data %>%
  dplyr::bind_rows() %>%
  dplyr::bind_rows(all_data %>%
                     dplyr::bind_rows() %>%
                     dplyr::mutate(`Study Area` = "UUSW Total")) %>%
  dplyr::filter(!is.na(`Study Area`)) %>%
  dplyr::mutate(`Study Area` = factor(`Study Area`, 
                                      levels = c("UUSW Total",
                                                 "CMV",
                                                 "NRG"))) %>%
  dplyr::group_by(Type,`Study Area`) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  tidyr::spread(Type, Count) %>%
  knitr::kable(caption = "Counts of radiocarbon and tree-ring dates in this study.",
               booktabs = TRUE,
               format.args = list(big.mark = ','))
```

@@

```{r date-densities, fig.height=9, fig.width=6.5, fig.cap='The number of radiocarbon and tree-ring dates by county in this study.', fig.show = 'hold', out.width = '50%'}
county_counts <-
  all_data %>%
  dplyr::bind_rows() %>%
  dplyr::group_by(Type, County) %>%
  dplyr::summarise(Dates = dplyr::n()) %>%
  dplyr::left_join(uusw_counties,
                   by = "County") %>%
  sf::st_as_sf()

theme_map <- function(base_size = 6.5,
                      base_family = ""){
  theme_bw(base_size = base_size,
           base_family = base_family) %+replace%
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          panel.background = element_blank(),
          panel.border = element_blank(),
          panel.grid = element_blank(),
          panel.grid.major = element_line(colour = 'transparent'),
          panel.grid.minor = element_line(colour = 'transparent'),
          legend.key = element_blank(),
          legend.position = "bottom",
          legend.background = element_blank(),
          legend.key.width = unit(0.15,"npc"),
          legend.text = element_text(size = rel(1)),
          plot.background = element_blank(),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

county_counts %>%
  dplyr::filter(Type == "Radiocarbon") %>%
  ggplot() +
  geom_sf(data = four_corners_states,
          fill = NA) +
  geom_sf(aes(fill = Dates),
          color = "black") +
  scale_fill_distiller(name = "Number of Radiocarbon Dates",
                       palette = "Greys",
                       direction = 1,
                       limits = c(0,500),
                       minor_breaks = seq(50,450,100),
                       breaks = seq(0,500,100),
                       expand = expand_scale(mult = 0, add = 0),
                       guide = guide_colorbar(title.position = "top")) +
  geom_sf(data = uusw_counties,
          fill = NA,
          color = "black") +
  theme_map(16)

county_counts %>%
  dplyr::filter(Type == "Tree-ring") %>%
  ggplot() +
  geom_sf(data = four_corners_states,
          fill = NA) +
  geom_sf(aes(fill = Dates),
          color = "black") +
  scale_fill_distiller(name = "Number of Tree-ring Dates",
                       palette = "Greys",
                       direction = 1,
                       limits = c(0,5000),
                       minor_breaks = seq(500,4500,1000),
                       breaks = seq(0,5000,1000),
                       expand = expand_scale(mult = 0, add = 0),
                       guide = guide_colorbar(title.position = "top")) +
  geom_sf(data = uusw_counties,
          fill = NA,
          color = "black") +
  theme_map(16)

```

```{r site-densities, fig.height=9, fig.width=6.5, fig.cap='The number of sites with radiocarbon and tree-ring dates by county in this study.', fig.show = 'hold', out.width = '50%'}
county_counts <-
  all_data %>%
  dplyr::bind_rows() %>%
  dplyr::select(Type, SiteID, County) %>%
  dplyr::distinct() %>%
  dplyr::group_by(Type, County) %>%
  dplyr::summarise(Dates = dplyr::n()) %>%
  dplyr::left_join(uusw_counties,
                   by = "County") %>%
  sf::st_as_sf()

theme_map <- function(base_size = 6.5,
                      base_family = ""){
  theme_bw(base_size = base_size,
           base_family = base_family) %+replace%
    theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          panel.background = element_blank(),
          panel.border = element_blank(),
          panel.grid = element_blank(),
          panel.grid.major = element_line(colour = 'transparent'),
          panel.grid.minor = element_line(colour = 'transparent'),
          legend.key = element_blank(),
          # legend.position = "top",
          # legend.justification = c(0,0),
          legend.position = "bottom",
          # legend.position = c(0.5,1),
          legend.background = element_blank(),
          legend.key.width = unit(0.15,"npc"),
          # legend.key.height = unit(0.15,"in"),
          legend.text = element_text(size = rel(1)),
          plot.background = element_blank(),
          plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

county_counts %>%
  dplyr::filter(Type == "Radiocarbon") %>%
  ggplot() +
  geom_sf(data = four_corners_states,
          fill = NA) +
  geom_sf(aes(fill = Dates),
          color = "black") +
  scale_fill_distiller(name = "Number of Sites with Radiocarbon Dates",
                       palette = "Greys",
                       direction = 1,
                       limits = c(0,80),
                       minor_breaks = seq(10,70,20),
                       breaks = seq(0,80,20),
                       expand = expand_scale(mult = 0, add = 0),
                       guide = guide_colorbar(title.position = "top")) +
  geom_sf(data = uusw_counties,
          fill = NA,
          color = "black") +
  theme_map(16)

county_counts %>%
  dplyr::filter(Type == "Tree-ring") %>%
  ggplot() +
  geom_sf(data = four_corners_states,
          fill = NA) +
  geom_sf(aes(fill = Dates),
          color = "black") +
  scale_fill_distiller(name = "Number of Sites with Tree-ring Dates",
                       palette = "Greys",
                       direction = 1,
                       limits = c(0,250),
                       minor_breaks = seq(25,225,50),
                       breaks = seq(0,250,50),
                       expand = expand_scale(mult = 0, add = 0),
                       guide = guide_colorbar(title.position = "top")) +
  geom_sf(data = uusw_counties,
          fill = NA,
          color = "black") +
  theme_map(16)

```

@@

all_data$Radiocarbon %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$Radiocarbon))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    rcarbon::calibrate(x = Date_BP,
                                                       errors = SD,
                                                       timeRange = timerange,
                                                       calMatrix = TRUE,
                                                       verbose = FALSE)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Calibrated,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        spdnormalised = TRUE,
                                        timeRange = timerange,
                                        runm = 21,
                                        verbose = FALSE))) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
                  purrr::map(function(x){
                    x$grid %>%
                      tibble::as_tibble() %$%
                      tibble::tibble(Type = "Radiocarbon Density",
                                     Date_BP = calBP,
                                     Density = PrDens / sum(PrDens, na.rm = T))
                  })
  ) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz")

radiocarbon_spd <- readr::read_rds(
  here::here("analysis/data/derived_data/radiocarbon_spd.rds")
)

```

@@

```{r treering-spds}
calibrate_tree_ring_dates <- function(x){
  out <- list(
    metadata = tibble::tibble(
      DateID = 1:length(x),
      CRA = x,
      Error = 0,
      Details = NA,
      CalCurve = "intcal13",
      ResOffsets = 0,
      ResErrors = 0,
      StartBP = timerange[1],
      EndBP = timerange[2],
      Normalised = TRUE,
      F14C = FALSE,
      CalEPS = 1e-05
    ),
    grids = NA,
    calmatrix = 
      tibble::tibble(index = 1:length(x),
                     year_bp = x,
                     value = TRUE) %>%
      tidyr::pivot_wider(names_from = index,
                         values_from = value) %>%
      dplyr::bind_rows(tibble::tibble(year_bp = timerange[1]:timerange[2])) %>%
      dplyr::filter(!duplicated(year_bp)) %>%
      dplyr::arrange(-year_bp) %>%
      dplyr::filter(year_bp >= 0) %>%
      tibble::column_to_rownames("year_bp") %>%
      as.matrix() %>%
      tidyr::replace_na(0) %>%
      magrittr::set_colnames(NULL)
  )

  attr(out,"class") <- c("CalDates", "list")

@@

all_data$`Tree-ring` %>%
  split(.,.$`Study Area`) %>%
  tibble::tibble(`Study Area` = names(.),
                 Dates = .) %>%
  dplyr::bind_rows(tibble::tibble(`Study Area` = "UUSW",
                                  Dates = list(all_data$`Tree-ring`))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(Calibrated = list(Dates %$%
                                    calibrate_tree_ring_dates(x = Date_BP)),
                Bins = list(rcarbon::binPrep(sites = Dates$SiteID,
                                             ages = Dates$Date_BP,
                                             h = 100)),
                SPD = list(rcarbon::spd(x = Calibrated,
                                        bins = Bins,
                                        spdnormalised = TRUE,
                                        timeRange = timerange,
                                        runm = 21,
                                        verbose = FALSE))) %>%

  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
                  purrr::map(function(x){
                    x$grid %>%
                      tibble::as_tibble() %$%
                      tibble::tibble(Type = "Tree-ring Density",
                                     Date_BP = calBP,
                                     Density = PrDens / sum(PrDens, na.rm = T))
                  })
  ) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/dendro_spd.rds"),
    compress = "gz")

dendro_spd <- readr::read_rds(
  here::here("analysis/data/derived_data/dendro_spd.rds")
)

```

@@

```{r c14-tr-pop, fig.height=6.5, fig.width=6.5, fig.cap='Radiocarbon and tree-ring density distributions in the UUSW, CMV, and NRG regions, and VEP population estimates for the CMV and NRG. Y-axis labels are omitted for simplicity; the maximum population in the lower panel is ~30,000 people, and the minimum is 0. Dates are given in AD.', fig.show = 'hold', out.width = '75%'}
all_density <- 
  dendro_spd %>%
  dplyr::select(`Study Area`,Density) %>%
  tidyr::unnest(cols = c(Density)) %>%
  dplyr::bind_rows(radiocarbon_spd %>%
                     dplyr::select(`Study Area`,Density) %>%
                     tidyr::unnest(cols = c(Density))) %>%
  dplyr::mutate(`Year AD` = 1950 - Date_BP) %>%
  dplyr::select(`Year AD`,
                Date_BP,
                `Study Area`,
                Type, 
                Density) %>%
  dplyr::rename(Value = Density)

vep_demography <- 
  robinson2020::vep_demography %>%
  dplyr::rowwise() %>%
  dplyr::mutate(`Year AD` = list((Start+1):End)) %>%
  dplyr::select(`Study Area`,
                `Year AD`,
                Population) %>%
  tidyr::unnest(`Year AD`) %>%
  dplyr::mutate(Type = "Population",
                Date_BP = 1950 - `Year AD`) %>%
  dplyr::rename(Value = Population)

>>>>

theme_map <- function(base_size = 6.5,
                      base_family = "") {
  theme_bw(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid = element_blank(),
      panel.grid.major = element_line(colour = "transparent"),
      panel.grid.minor = element_line(colour = "transparent"),
      legend.key = element_blank(),
      legend.position = "top",
      plot.background = element_blank(),
      plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

ggplot() +
  geom_sf(data = four_corners_states) +
  geom_raster(
    data = robinson2020::swus_hillshade_500m,
    mapping = aes(
      x = x,
      y = y,
      alpha = ID
    ),
    na.rm = TRUE
  ) +
  scale_alpha(
    range = c(0.8, 0),
    na.value = 0,
    limits = c(0, 255),
    guide = "none"
  ) +
  geom_sf(
    data = four_corners_states,
    fill = NA
  ) +
  geom_sf(
    data = uusw_counties %>%
      sf::st_union() %>%
      rmapshaper::ms_filter_islands(),
    fill = NA,
    color = "black"
  ) +
  geom_sf(
    data = uusw_boundary,
    fill = NA,
    linetype = 2,
    color = "white"
  ) +
  geom_sf(
    data = vep_study_areas,
    fill = NA,
    color = "white",
    size = 1
  ) +
  geom_sf_text(
    data = vep_study_areas,
    aes(label = `Study Area`),
    size = 3
  ) +
  theme_map(16)
```

@@

```{r prepare-radiocarbon-data}
# Read in the radiocarbon database
if (!file.exists(
  here::here("analysis/data/derived_data/radiocarbon_data.csv")
)) {
  readr::read_csv(
    here::here("analysis/data/raw_data/0403_sbox_clean_sd300.csv")
  ) %>%
    # Convert to spatial object
    sf::st_as_sf(
      coords = c("Long", "Lat"),
      crs = 4326
    ) %>%
    sf::st_transform(four_corners_lcc_proj) %>%
    # Select salient columns
    dplyr::select(
      labnumber,
      date,
      sd,
      Site,
      SiteID
    ) %>%
    dplyr::rename(
      Lab_Number = labnumber,
      Date_BP = date,
      SD = sd,
      Site_Number = Site
    ) %>%
    sf::st_intersection(uusw_counties) %>%
    sf::st_drop_geometry() %>%
    dplyr::mutate(
      Type = "Radiocarbon",
      SiteID = SiteID %>% stringr::str_c("Radiocarbon - ", .)
    ) %>%
    dplyr::select(
      Type,
      SiteID,
      Site_Number,
      County,
      Lab_Number,
      Date_BP,
      SD
    ) %>%
    # Write the output to a new file (csv)
    readr::write_csv(
      here::here("analysis/data/derived_data/radiocarbon_data.csv")
    )
}

radiocarbon_data <-
  readr::read_csv(
    here::here("analysis/data/derived_data/radiocarbon_data.csv")
  )
```

@@

if (!file.exists(
  here::here("analysis/data/derived_data/dendro_data.csv")
)) {
  readr::read_csv(
    here::here("analysis/data/raw_data/BOCINSKY_ET_AL_2015_TREE_RINGS.csv")
  ) %>%
    # Convert to spatial object
    sf::st_as_sf(
      coords = c("LON", "LAT"),
      crs = 4326
    ) %>%
    sf::st_transform(four_corners_lcc_proj) %>%
    # Convert dates to BP
    dplyr::mutate(
      Date_BP = 1950 - Outer_Date_AD,
      Type = ifelse(C_Level %in% c(2, 3),
        "Cutting or Near-cutting",
        "Non-cutting"
      )
    ) %>%
    # drop C_Level
    dplyr::select(
      Seq_Number,
      Site_Number,
      Lab_Number,
      Type,
      Date_BP
    ) %>%
    dplyr::rename(SiteID = Seq_Number) %>%
    sf::st_intersection(uusw_counties) %>%
    sf::st_drop_geometry() %>%
    dplyr::mutate(
      Type = Type %>% stringr::str_c("Dendro - ", .),
      SiteID = SiteID %>% stringr::str_c("Dendro - ", .),
      SD = 0
    ) %>%
    dplyr::filter(Type == "Dendro - Cutting or Near-cutting") %>%
    dplyr::mutate(Type = "Tree-ring") %>%
    dplyr::select(
      Type,
      SiteID,
      Site_Number,
      County,
      Lab_Number,
      Date_BP,
      SD
    ) %>%
    # Write the output to a new file (csv)
    readr::write_csv(
      here::here("analysis/data/derived_data/dendro_data.csv")
    )
}

dendro_data <-
  readr::read_csv(
    here::here("analysis/data/derived_data/dendro_data.csv")
  )
```

@@

```{r}
all_data <- list(
  Radiocarbon = radiocarbon_data,
  `Tree-ring` = dendro_data
) %>%
  purrr::map(~ dplyr::mutate(.x,
    `Study Area` = ifelse(County == "Montezuma, CO",
      "CMV",
      NA
    ),
    `Study Area` = ifelse(County %in% c(
      "Los Alamos, NM",
      "Rio Arriba, NM",
      "Sandoval, NM",
      "Santa Fe, NM",
      "Taos, NM"
    ),
    "NRG",
    `Study Area`
    )
  ))
```

@@

```{r data-summary}
all_data %>%
  dplyr::bind_rows() %>%
  dplyr::bind_rows(all_data %>%
    dplyr::bind_rows() %>%
    dplyr::mutate(`Study Area` = "UUSW Total")) %>%
  dplyr::filter(!is.na(`Study Area`)) %>%
  dplyr::mutate(`Study Area` = factor(`Study Area`,
    levels = c(
      "UUSW Total",
      "CMV",
      "NRG"
    )
  )) %>%
  dplyr::group_by(Type, `Study Area`) %>%
  dplyr::summarise(Count = dplyr::n()) %>%
  tidyr::spread(Type, Count) %>%
  knitr::kable(
    caption = "Counts of radiocarbon and tree-ring dates in this study.",
    booktabs = TRUE,
    format.args = list(big.mark = ",")
  )
```

@@

```{r date-densities, fig.height=9, fig.width=6.5, fig.cap='The number of radiocarbon and tree-ring dates by county in this study.', fig.show = 'hold', out.width = '50%'}
county_counts <-
  all_data %>%
  dplyr::bind_rows() %>%
  dplyr::group_by(Type, County) %>%
  dplyr::summarise(Dates = dplyr::n()) %>%
  dplyr::left_join(uusw_counties,
    by = "County"
  ) %>%
  sf::st_as_sf()

theme_map <- function(base_size = 6.5,
                      base_family = "") {
  theme_bw(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid = element_blank(),
      panel.grid.major = element_line(colour = "transparent"),
      panel.grid.minor = element_line(colour = "transparent"),
      legend.key = element_blank(),
      legend.position = "bottom",
      legend.background = element_blank(),
      legend.key.width = unit(0.15, "npc"),
      legend.text = element_text(size = rel(1)),
      plot.background = element_blank(),
      plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

county_counts %>%
  dplyr::filter(Type == "Radiocarbon") %>%
  ggplot() +
  geom_sf(
    data = four_corners_states,
    fill = NA
  ) +
  geom_sf(aes(fill = Dates),
    color = "black"
  ) +
  scale_fill_distiller(
    name = "Number of Radiocarbon Dates",
    palette = "Greys",
    direction = 1,
    limits = c(0, 500),
    minor_breaks = seq(50, 450, 100),
    breaks = seq(0, 500, 100),
    expand = expand_scale(mult = 0, add = 0),
    guide = guide_colorbar(title.position = "top")
  ) +
  geom_sf(
    data = uusw_counties,
    fill = NA,
    color = "black"
  ) +
  theme_map(16)

county_counts %>%
  dplyr::filter(Type == "Tree-ring") %>%
  ggplot() +
  geom_sf(
    data = four_corners_states,
    fill = NA
  ) +
  geom_sf(aes(fill = Dates),
    color = "black"
  ) +
  scale_fill_distiller(
    name = "Number of Tree-ring Dates",
    palette = "Greys",
    direction = 1,
    limits = c(0, 5000),
    minor_breaks = seq(500, 4500, 1000),
    breaks = seq(0, 5000, 1000),
    expand = expand_scale(mult = 0, add = 0),
    guide = guide_colorbar(title.position = "top")
  ) +
  geom_sf(
    data = uusw_counties,
    fill = NA,
    color = "black"
  ) +
  theme_map(16)
```

```{r site-densities, fig.height=9, fig.width=6.5, fig.cap='The number of sites with radiocarbon and tree-ring dates by county in this study.', fig.show = 'hold', out.width = '50%'}
county_counts <-
  all_data %>%
  dplyr::bind_rows() %>%
  dplyr::select(Type, SiteID, County) %>%
  dplyr::distinct() %>%
  dplyr::group_by(Type, County) %>%
  dplyr::summarise(Dates = dplyr::n()) %>%
  dplyr::left_join(uusw_counties,
    by = "County"
  ) %>%
  sf::st_as_sf()

theme_map <- function(base_size = 6.5,
                      base_family = "") {
  theme_bw(
    base_size = base_size,
    base_family = base_family
  ) %+replace%
    theme(
      axis.line = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      axis.title = element_blank(),
      panel.background = element_blank(),
      panel.border = element_blank(),
      panel.grid = element_blank(),
      panel.grid.major = element_line(colour = "transparent"),
      panel.grid.minor = element_line(colour = "transparent"),
      legend.key = element_blank(),
      legend.position = "bottom",
      legend.background = element_blank(),
      legend.key.width = unit(0.15, "npc"),
      legend.text = element_text(size = rel(1)),
      plot.background = element_blank(),
      plot.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "npc")
    )
}

county_counts %>%
  dplyr::filter(Type == "Radiocarbon") %>%
  ggplot() +
  geom_sf(
    data = four_corners_states,
    fill = NA
  ) +
  geom_sf(aes(fill = Dates),
    color = "black"
  ) +
  scale_fill_distiller(
    name = "Number of Sites with Radiocarbon Dates",
    palette = "Greys",
    direction = 1,
    limits = c(0, 80),
    minor_breaks = seq(10, 70, 20),
    breaks = seq(0, 80, 20),
    expand = expand_scale(mult = 0, add = 0),
    guide = guide_colorbar(title.position = "top")
  ) +
  geom_sf(
    data = uusw_counties,
    fill = NA,
    color = "black"
  ) +
  theme_map(16)

county_counts %>%
  dplyr::filter(Type == "Tree-ring") %>%
  ggplot() +
  geom_sf(
    data = four_corners_states,
    fill = NA
  ) +
  geom_sf(aes(fill = Dates),
    color = "black"
  ) +
  scale_fill_distiller(
    name = "Number of Sites with Tree-ring Dates",
    palette = "Greys",
    direction = 1,
    limits = c(0, 250),
    minor_breaks = seq(25, 225, 50),
    breaks = seq(0, 250, 50),
    expand = expand_scale(mult = 0, add = 0),
    guide = guide_colorbar(title.position = "top")
  ) +
  geom_sf(
    data = uusw_counties,
    fill = NA,
    color = "black"
  ) +
  theme_map(16)
```

@@

all_data$Radiocarbon %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$Radiocarbon)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      rcarbon::calibrate(
        x = Date_BP,
        errors = SD,
        timeRange = timerange,
        calMatrix = TRUE,
        verbose = FALSE
      )),
    Bins = list(rcarbon::binPrep(
      sites = Dates$SiteID,
      ages = Calibrated,
      h = 100
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      spdnormalised = TRUE,
      timeRange = timerange,
      runm = 21,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Radiocarbon Density",
          Date_BP = calBP,
          Density = PrDens / sum(PrDens, na.rm = T)
        )
    })) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/radiocarbon_spd.rds"),
    compress = "gz"
  )

radiocarbon_spd <- readr::read_rds(
  here::here("analysis/data/derived_data/radiocarbon_spd.rds")
)
```

@@

```{r treering-spds}
calibrate_tree_ring_dates <- function(x) {
  out <- list(
    metadata = tibble::tibble(
      DateID = 1:length(x),
      CRA = x,
      Error = 0,
      Details = NA,
      CalCurve = "intcal13",
      ResOffsets = 0,
      ResErrors = 0,
      StartBP = timerange[1],
      EndBP = timerange[2],
      Normalised = TRUE,
      F14C = FALSE,
      CalEPS = 1e-05
    ),
    grids = NA,
    calmatrix =
      tibble::tibble(
        index = 1:length(x),
        year_bp = x,
        value = TRUE
      ) %>%
        tidyr::pivot_wider(
          names_from = index,
          values_from = value
        ) %>%
        dplyr::bind_rows(tibble::tibble(year_bp = timerange[1]:timerange[2])) %>%
        dplyr::filter(!duplicated(year_bp)) %>%
        dplyr::arrange(-year_bp) %>%
        dplyr::filter(year_bp >= 0) %>%
        tibble::column_to_rownames("year_bp") %>%
        as.matrix() %>%
        tidyr::replace_na(0) %>%
        magrittr::set_colnames(NULL)
  )

  attr(out, "class") <- c("CalDates", "list")

@@

all_data$`Tree-ring` %>%
  split(., .$`Study Area`) %>%
  tibble::tibble(
    `Study Area` = names(.),
    Dates = .
  ) %>%
  dplyr::bind_rows(tibble::tibble(
    `Study Area` = "UUSW",
    Dates = list(all_data$`Tree-ring`)
  )) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    Calibrated = list(Dates %$%
      calibrate_tree_ring_dates(x = Date_BP)),
    Bins = list(rcarbon::binPrep(
      sites = Dates$SiteID,
      ages = Dates$Date_BP,
      h = 100
    )),
    SPD = list(rcarbon::spd(
      x = Calibrated,
      bins = Bins,
      spdnormalised = TRUE,
      timeRange = timerange,
      runm = 21,
      verbose = FALSE
    ))
  ) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Density = SPD %>%
    purrr::map(function(x) {
      x$grid %>%
        tibble::as_tibble() %$%
        tibble::tibble(
          Type = "Tree-ring Density",
          Date_BP = calBP,
          Density = PrDens / sum(PrDens, na.rm = T)
        )
    })) %>%
  readr::write_rds(
    here::here("analysis/data/derived_data/dendro_spd.rds"),
    compress = "gz"
  )

dendro_spd <- readr::read_rds(
  here::here("analysis/data/derived_data/dendro_spd.rds")
)
```

@@

```{r c14-tr-pop, fig.height=6.5, fig.width=6.5, fig.cap='Radiocarbon and tree-ring density distributions in the UUSW, CMV, and NRG regions, and VEP population estimates for the CMV and NRG. Y-axis labels are omitted for simplicity; the maximum population in the lower panel is ~30,000 people, and the minimum is 0. Dates are given in AD.', fig.show = 'hold', out.width = '75%'}
all_density <-
  dendro_spd %>%
  dplyr::select(`Study Area`, Density) %>%
  tidyr::unnest(cols = c(Density)) %>%
  dplyr::bind_rows(radiocarbon_spd %>%
    dplyr::select(`Study Area`, Density) %>%
    tidyr::unnest(cols = c(Density))) %>%
  dplyr::mutate(`Year AD` = 1950 - Date_BP) %>%
  dplyr::select(
    `Year AD`,
    Date_BP,
    `Study Area`,
    Type,
    Density
  ) %>%
  dplyr::rename(Value = Density)

vep_demography <-
  robinson2020::vep_demography %>%
  dplyr::rowwise() %>%
  dplyr::mutate(`Year AD` = list((Start + 1):End)) %>%
  dplyr::select(
    `Study Area`,
    `Year AD`,
    Population
  ) %>%
  tidyr::unnest(`Year AD`) %>%
  dplyr::mutate(
    Type = "Population",
    Date_BP = 1950 - `Year AD`
  ) %>%
  dplyr::rename(Value = Population)

all_density %>%
  dplyr::bind_rows(vep_demography) %>%
  dplyr::mutate(
    Type = factor(Type,
      levels = c(
        "Radiocarbon Density",
        "Tree-ring Density",
        "Population"
      ),
      ordered = TRUE
    ),
    `Study Area` = factor(`Study Area`,
      levels = c(
        `Upland US Southwest` = "UUSW",
        `Central Mesa Verde` = "CMV",
        `Northern Rio Grande` = "NRG"
      )
    )
  ) %>%
  ggplot(aes(x = `Year AD`)) +
  geom_line(aes(
    y = Value,
    color = `Study Area`,
    size = `Study Area`
  )) +
  scale_color_manual(values = c(
    "UUSW" = "black",
    "CMV" = "gray50",
    "NRG" = "gray80"
  )) +
  scale_size_manual(values = c(
    "UUSW" = 1.25,
    "CMV" = 0.75,
    "NRG" = 0.75
  )) +
  # coord_cartesian(xlim = c(600,1600)) +
  xlab("Year AD") +
  scale_x_continuous(breaks = c(1, seq(200, 2000, 200))) +
  theme_minimal(18) +
  theme(
    legend.title = element_blank(),
    legend.justification = c(0, 1),
    legend.position = "bottom",
    axis.title.x = element_text(size = 12),
    axis.text.x = element_text(size = 10),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.margin = unit(c(0.25, 0.5, 0.25, 0.25), "in")
  ) +
  facet_wrap("Type",
    nrow = 3,
    scales = "free_y"
  )
```
